Draft: Notes on the revised Hidden weight function

### 2024-10-07 RevisedHiddenWeight.tex
Proving bound 1 + n * 2^{3n/4} for |W_f(a)|.

\begin{Prop}\label{prop:generating_fct}
	It holds that $\sum_{k\in\Z}\Dnka nka z^k= P_a(z)$.
\end{Prop}

@@

\begin{Prop}\label{prop:symmetry}
	For every integer $k$, we have:
	\[
	\Dnka n{n-k}a=(-1)^{p+q+k}\Dnka nka.
	\]
\end{Prop}

@@

\begin{proof}
	By Property \ref{prop:symmetry}, we have $\Dnka n{n/2}a=(-1)^{p+q+n/2}\Dnka n{n/2}a=-\Dnka n{n/2}a$ because $p+q+n/2\equiv n/2-p-q\equiv r\equiv 1\bmod 2$.
\end{proof}

Another immediate case for zero coefficients is the following:
\begin{Prop}

    If $p=q$ and $k$ is odd, then $\Dnka nka=0=\Dnka n{n-k}a$.
\end{Prop}

\begin{proof}
    If $p=q$, we have $P_a(z)=(z^4-6z^2+1)^p\cdot(z^2+1)^r$, which can be seen as a polynomial in $z^2$. Thus, the coefficient $\Dnka nka$ of $z^k$ for odd $k$ must be zero in $P_a(z)$. It follows that also $\Dnka n{n-k}a=0$ by Property \ref{prop:symmetry}.
\end{proof}

\begin{remark}
    We can also use Property \ref{prop:generating_fct} to obtain $\Dnka nka$ through differentiation:
    \[
        k!\cdot\Dnka nka=\frac{\dd^k}{\dd^k z}P_a(z)\vert_{z=0}.
    \]
    Using this formula, we can for instance use a computer algebra system to deduce $\Dnka nka$ for small values of $k$, see Table \ref{tab:Dnka_small_k}. Note that even though the expression for $\Dnka n3a$ involves a division by $3$, its evaluation at specific values for $(p,q,r)$ will always yield integers. Indeed, either $p-q\equiv 0\bmod 3$, or $(p-q)^2\equiv 1\bmod 3$, in which case $2(p-q)^2+7\equiv 0\bmod 3$.

@@

\begin{Prop}\mbox{}
    \begin{enumerate}
        \item If $n=16m$ and $\{p,q\}=\{m+\ell_s,m+\ell_{s+1}\}$ for integers $m,s\geq 0$, where $\ell_s=s(s-1)$, then $\Dnka n2a=0=\Dnka n{k-2}a$.
        \item If $n=16m+2$ and $\{p,q\}=\{m+\ell_s,m+\ell_{s+1}\}$ for integers $m,s\geq 0$, where $\ell_s=\left(6s^2+6s+(-1)^s(2s+1)-1\right)/8$, then $\Dnka n3a=0=\Dnka n{k-3}a$.
        \item If $n=16m+4$ and $\{p,q\}=\{m+\ell_s,m+\ell_{s+1}\}$ for integers $m,s\geq 0$, where $\ell_s=s^2$, then $\Dnka n2a=0=\Dnka n{k-2}a$.
        \item If $n=16m+6$ and $\{p,q\}=\{m+\ell_s,m+\ell_{s+1}\}$ for integers $m,s\geq 0$, where $\ell_s=\left(6s^2+6s-(-1)^s(2s+1)+1\right)/8$, then $\Dnka n3a=0=\Dnka n{k-3}a$.
    \end{enumerate}
\end{Prop}

@@

Using Property \ref{prop:generating_fct}, we can also deduce some sums involving the $\Dnka nka$:

\begin{Prop}\mbox{}
    \begin{enumerate}
        \item The sum and alternating sum over $k$ of the $\Dnka nka$ are, respectively:
        \begin{gather*}
            \sum_{k=0}^n\Dnka nka=(-1)^{q}2^{n/2},\\
            \sum_{k=0}^n(-1)^k\Dnka nka=(-1)^{p}2^{n/2}.
        \end{gather*}
        \item The sum over the even and the odd $k$ of the the $\Dnka nka$ are, respectively:
        \begin{gather*}
            \sum_{k=0}^{n/2}\Dnka n{2k}a=2^{n/2-1}\big((-1)^{q}+(-1)^{p}\big),\\
            \sum_{k=0}^{n/2-1}\Dnka n{2k+1}a=2^{n/2-1}\big((-1)^{q}-(-1)^{p}\big).
        \end{gather*}
    \end{enumerate}

\end{Prop}

@@

One can also prove the following:

\begin{Prop}
    For a fixed integer $k\geq 1$ and sufficiently large $n\geq b(k)$, it holds that:
    \[
        \max_{a\in\F_2^n}|\Dnka nka|=|\Dnka nk{0_n}|=\left(-\frac12\right)^k\sum_{i=\lceil k/2\rceil}^{n/2}(-4)^i\binom{n/2}{i}\binom{i}{k-i}.
    \]
\end{Prop}

\begin{remark}
    In general, it also appears that the vector $a\in\F_2^n$ for which $|\Dnka nka|$ is maximal always satisfies $r(a)=0$.
\end{remark}

@@

\begin{Prop}
    For any real number $R>0$, we have the following bound:
    \[
        |\Dnka nka|\leq\frac{(M_R)^{n/2}}{R^k},
    \]
    where $M_R$ is the maximum of $\left|z^2-2z-1\right|$ on the complex disk $|z|=R$. Explicitly, $M_R$ can be expressed as follows:
    \[
        M_R=
        \begin{cases}
            -R^2+2R+1&\text{if $R\leq\sqrt 2-1$}\\
            \sqrt 2\left(R^2+1\right)&\text{if $\sqrt 2-1\leq R\leq\sqrt 2+1$}\\
            R^2+2R-1&\text{if $R\geq\sqrt 2+1$}\\
        \end{cases}.
    \]
\end{Prop}

@@

\begin{remark}
    We can even go ahead and choose $R=R_{n,k}$ in terms of $n$ and $k$, in such a way that the upper bound of $B_{R,n,k}=(M_R)^{n/2}/R^k$ is minimized. To do this, consider any $1\leq k\leq n-1$. We can actually assume that $k\leq n/2$: if $k>n/2$, replace it by $n-k$ (recall that $|\Dnka nka|=|\Dnka n{n-k}a|$ by Property \ref{prop:symmetry}). One can show that $B_{R,n,k}$ is minimal for the following choice of $R=R_{n,k}$:
    \[
        R_{n,k}=\begin{cases}
            \frac{n-2k-\sqrt{2(n-2k)^2-n^2}}{2(n-k)}&\text{if $k\leq\frac{2-\sqrt 2}{4}\cdot n$}\\
            \sqrt{\frac{k}{n-k}}&\text{if $k\geq\frac{2-\sqrt 2}{4}\cdot n$}
        \end{cases}.
    \]
    Figure \ref{fig:cauchy_estimate} compares for $n=1\,000$ the two upper bounds we have for $|\Dnka nka|$ on a logarithmic scale: the binomial coefficient $\binom nk$ in red, and the bound $B_{R,n,k}$ with the optimally chosen $R=R_{n,k}$ in blue. The $x$-axis represents the values for $k$, while the $y$-axis represents the binary logarithm of the two bounds. In case $|k-n/2|\leq\sqrt{2}/4\cdot n$, it can be observed that the Cauchy estimate yields particularly good bounds compared to the binomial coefficients. In fact, in this case, we get:
    \[
        |\Dnka nka|\leq 2^{n/4}\sqrt{\frac{n^n}{(n-k)^{n-k}\cdot k^k}}.
    \]
    \begin{figure}
        \centering
        \includegraphics[width=7cm]{cauchy_estimate.pdf}
        \caption{Cauchy estimate in blue vs.\ binomial coefficients in red}
        \label{fig:cauchy_estimate}
    \end{figure}
\end{remark}

\begin{Prop}\label{prop:bound_sum_D_nka}
    For every $1\leq k\leq n$, fix a Boolean vector $a\ui k\in\F_2^n$. Then we have the following bound:
    \[
        \sum_{k=1}^n\left|\Dnka nk{a\ui k}\right|\leq n\cdot 2^{3n/4}.
    \]
\end{Prop}

\begin{proof}
    Simply observe that $2^{n/4}\sqrt{\frac{n^n}{(n-k)^{n-k}\cdot k^k}}$ is maximal for $k=n/2$, in which case the expression reduces to $2^{3n/4}$.
\end{proof}

\subsection{Bounding the Walsh transform of $f$}

From Property \ref{prop:bound_sum_D_nka}, we can in particular deduce that $|\wt f(a)|\leq n\cdot 2^{3n/4}+1$ for every Boolean vector $a\in\F_2^n$. This is a very pessimistic bound. In this section, we will discuss how a tighter bound on $|\wt f(a)|$ can be obtained in polynomial time in $n$.

@@

Consider now the complexity involved in computing the values of $\mathsf{D}_{n,k}^{p,q,r}$ for all $0 \leq k \leq n$ and for all triplets $(p,q,r)$ where $p+q+r=n/2$. We claim that this computation has a complexity of $\mathcal{O}(n^3 \log n)$. Towards this, according to Property \ref{prop:generating_fct}, it is sufficient to expand the polynomials $(-z^2 + 2z + 1)^p \cdot (-z^2 - 2z + 1)^q \cdot (z^2 + 1)^r$ for all triplets $(p,q,r)$ such that $p+q+r=n/2$.

@@

Figure \ref{fig:walsh_bound_comparison} compares our experimental bound $B_n$ to the theoretical bound $n\cdot 2^{3n/4}+1$ and the binomial coefficient $\binom{n}{n/2}$, on a logarithmic scale. The first two and the last one have an asymptotic slope of $3/4$ and $1$, respectively.

\begin{figure}
    \centering
    \includegraphics[width=9cm]{comparison_walsh_bound.pdf}
    \caption{On a logarithmic scale, our experimental bound $B_n$ in terms of $n$ in red, our theoretical bound $n\cdot 2^{3n/4}+1$ from Property \ref{prop:bound_sum_D_nka} in green, and the bound $\binom{n}{n/2}$ in blue.}
    \label{fig:walsh_bound_comparison}
\end{figure}

@@

\begin{proposition}\label{prop:dtn}
	Let $t, n\in \N$ with $n$ even and $t\le n/2$, the following holds on $d_{t,n}$ for all $k\in [0,n]$:
	\[\wtk{d_{t,n}}{k}(a)=\sum_{\ell=0}^{2t} \Dnka{2t}{\ell}{b} \kraw{k-\ell}{\w(c)}{n-2t},\]
	where $b$ denotes the first $2t$ elements of $a$ and $c$ the last $n-2t$ elements.

>>>>

\begin{proposition}\label{proposition:generating_fct}
	It holds that $\sum_{k\in\Z}\Dnka nka z^k= P_a(z)$.
\end{proposition}

@@

\begin{proposition}\label{proposition:symmetry}
	For every integer $k$, we have:
	\[
	\Dnka n{n-k}a=(-1)^{p+q+k}\Dnka nka.
	\]
\end{proposition}

@@

\begin{proof}
	By Proposition \ref{proposition:symmetry}, we have $\Dnka n{n/2}a=(-1)^{p+q+n/2}\Dnka n{n/2}a=-\Dnka n{n/2}a$ because $p+q+n/2\equiv n/2-p-q\equiv r\equiv 1\bmod 2$.
\end{proof}

Another immediate case for zero coefficients is the following:
\begin{proposition}
    If $p=q$ and $k$ is odd, then $\Dnka nka=0=\Dnka n{n-k}a$.
\end{proposition}

\begin{proof}
    If $p=q$, we have $P_a(z)=(z^4-6z^2+1)^p\cdot(z^2+1)^r$, which can be seen as a polynomial in $z^2$. Thus, the coefficient $\Dnka nka$ of $z^k$ for odd $k$ must be zero in $P_a(z)$. It follows that also $\Dnka n{n-k}a=0$ by Proposition \ref{proposition:symmetry}.
\end{proof}

\begin{remark}\label{remark:D_nka_differentiation}
    We can also use Proposition \ref{proposition:generating_fct} to obtain $\Dnka nka$ through differentiation:
    \[
        k!\cdot\Dnka nka=\frac{\dd^k}{\dd^k z}P_a(z)\vert_{z=0}.
    \]
    Using this formula, we can for instance use a computer algebra system to deduce $\Dnka nka$ for small values of $k$, see Table \ref{tab:Dnka_small_k}. Note that even though the expression for $\Dnka n3a$ involves a division by $3$, its evaluation at specific values for $(p,q,r)$ will always yield integers. Indeed, either $p-q\equiv 0\bmod 3$, or $(p-q)^2\equiv 1\bmod 3$, in which case $2(p-q)^2+7\equiv 0\bmod 3$.

@@

\begin{proposition}\mbox{}
    \begin{enumerate}
        \item If $n=16m$ and $\{p,q\}=\{m+\ell_s,m+\ell_{s+1}\}$ for integers $m,s\geq 0$, where $\ell_s=s(s-1)$, then $\Dnka n2a=0=\Dnka n{k-2}a$.
        \item If $n=16m+2$ and $\{p,q\}=\{m+\ell_s,m+\ell_{s+1}\}$ for integers $m,s\geq 0$, where $\ell_s=\left(6s^2+6s+(-1)^s(2s+1)-1\right)/8$, then $\Dnka n3a=0=\Dnka n{k-3}a$.
        \item If $n=16m+4$ and $\{p,q\}=\{m+\ell_s,m+\ell_{s+1}\}$ for integers $m,s\geq 0$, where $\ell_s=s^2$, then $\Dnka n2a=0=\Dnka n{k-2}a$.
        \item If $n=16m+6$ and $\{p,q\}=\{m+\ell_s,m+\ell_{s+1}\}$ for integers $m,s\geq 0$, where $\ell_s=\left(6s^2+6s-(-1)^s(2s+1)+1\right)/8$, then $\Dnka n3a=0=\Dnka n{k-3}a$.
    \end{enumerate}
\end{proposition}

@@

Using Proposition \ref{proposition:generating_fct}, we can also deduce some sums involving the $\Dnka nka$:

\begin{proposition}\mbox{}
    \begin{enumerate}
        \item The sum and alternating sum over $k$ of the $\Dnka nka$ are, respectively:
        \begin{gather*}
            \sum_{k=0}^n\Dnka nka=(-1)^{q}2^{n/2},\\
            \sum_{k=0}^n(-1)^k\Dnka nka=(-1)^{p}2^{n/2}.
        \end{gather*}
        \item The sum over the even and the odd $k$ of the the $\Dnka nka$ are, respectively:
        \begin{gather*}
            \sum_{k=0}^{n/2}\Dnka n{2k}a=2^{n/2-1}\big((-1)^{q}+(-1)^{p}\big),\\
            \sum_{k=0}^{n/2-1}\Dnka n{2k+1}a=2^{n/2-1}\big((-1)^{q}-(-1)^{p}\big).
        \end{gather*}
    \end{enumerate}

\end{proposition}

@@

\subsection{Bounding the Walsh transform of $f$ using the Cauchy estimate}

Recall the Cauchy estimate on holomorphic functions:

\begin{Prop}[\cite{rudin1987}, Theorem 10.26]\label{property:cauchy_estimate}
    Let $w$ be a complex number, and let $R>0$ be some radius. If $f$ is a holomorphic function defined on an open set containing every complex number $z$ satisfying $|z-w|<R$, then for every integer $k\geq 0$, the $k$-th derivative of $f$ evaluated at $w$ can be bounded, in absolute value, as follows:
    \[
        \left|\frac{\dd^k}{\dd^k z}f(w)\right|\leq\frac{k!\cdot M_R}{R^k},
    \]
    where $M_R$ is the maximum of $|f(z)|$ on the circle $|z-w|=R$.
\end{Prop}

Applying this, we get:

\begin{theorem}\label{theorem:bound_D_nka}
    For every $a\in\F_2^n$ and every integer $k\geq 0$, it holds that:
    \[
        |\Dnka nka|\leq 2^{3n/4}.
    \]
\end{theorem}

\begin{proof}
    Recall from Remark \ref{remark:D_nka_differentiation} that $k!\cdot\Dnka nka=\frac{\dd^k}{\dd^k z}P_a(z)\vert_{z=0}$. Choosing $f=P_a$, $w=0$ and $R=1$ in Property \ref{property:cauchy_estimate} then gives:
    \[
        |\Dnka nka|=\frac{1}{k!}\left|\frac{\dd^k}{\dd^k z}P_a(z)\vert_{z=0}\right|\leq M,
    \]
    where $M$ is the maximum of $|P_a(z)|$ on the circle $|z|=1$. We claim, and prove below, that the maximum of $\left|-z^2+2z+1\right|$ on $|z|=1$ is $2\sqrt 2$. This implies that the maximum of $\left|-z^2-2z+1\right|$ on $|z|=1$ is $2\sqrt 2$ as well. Since the maximum of $\left|z^2+1\right|$ on $|z|=1$ is $2<2\sqrt 2$, it follows that $|\Dnka nka|\leq M\leq (2\sqrt 2)^{p+q}2^r\leq (2\sqrt 2)^{n/2}$.

    It remains to prove that the maximum of $\left|-z^2+2z+1\right|^2$ on $|z|=1$ equals $8$. Using that $|w|^2=w\overline w$ for every complex number $w$, we obtain for $|z|=1$:
    \[
        \left|-z^2+2z+1\right|^2=\left(z^2-2z-1\right)\left(\overline z^2-2\overline z-1\right)=6-\left(z^2+\overline z^2\right)=6-2\mathrm{Re}\left(z^2\right),
    \]
    where $\mathrm{Re}\left(z^2\right)$ denotes the real part of $z^2$. Since the minimum of $\mathrm{Re}\left(z^2\right)$ on $|z|=1$ is $-1$, we conclude that the maximum of $\left|-z^2+2z+1\right|^2$ on $|z|=1$ is $6-2(-1)=8$.
\end{proof}

As a consequence, we have:

\begin{Corollary}\label{cor:bound_walsh_cauchy}
    For every $a\in\F_2^n$, it holds that:
    \[
        |\wt f(a)|\leq 1+n\cdot 2^{3n/4}.
    \]
\end{Corollary}

\begin{proof}
    We have previously established that $\wt{f}(a)=1+\sum_{k=1}^n\Dnka{n}{k}{b+\pi(e_k)}$. Then, applying Proposition \ref{theorem:bound_D_nka} yields:
    \[
        |\wt{f}(a)|\leq 1+\sum_{k=1}^n\left|\Dnka{n}{k}{b+\pi(e_k)}\right|\leq 1+ n\cdot 2^{3n/4}.\qedhere
    \]
\end{proof}

\mycomment{One can also prove the following:

\begin{proposition}
    For a fixed integer $k\geq 1$ and sufficiently large $n\geq b(k)$, it holds that:
    \[
        \max_{a\in\F_2^n}|\Dnka nka|=|\Dnka nk{0_n}|=\left(-\frac12\right)^k\sum_{i=\lceil k/2\rceil}^{n/2}(-4)^i\binom{n/2}{i}\binom{i}{k-i}.
    \]
\end{proposition}

@@

\begin{proposition}
    For any real number $R>0$, we have the following bound:
    \[
        |\Dnka nka|\leq\frac{(M_R)^{n/2}}{R^k},
    \]
    where $M_R$ is the maximum of $\left|z^2-2z-1\right|$ on the complex disk $|z|=R$. Explicitly, $M_R$ can be expressed as follows:
    \[
        M_R=
        \begin{cases}
            -R^2+2R+1&\text{if $R\leq\sqrt 2-1$}\\
            \sqrt 2\left(R^2+1\right)&\text{if $\sqrt 2-1\leq R\leq\sqrt 2+1$}\\
            R^2+2R-1&\text{if $R\geq\sqrt 2+1$}\\
        \end{cases}.
    \]
\end{proposition}

@@

\begin{remark}
    We can even go ahead and choose $R=R_{n,k}$ in terms of $n$ and $k$, in such a way that the upper bound of $B_{R,n,k}=(M_R)^{n/2}/R^k$ is minimized. To do this, consider any $1\leq k\leq n-1$. We can actually assume that $k\leq n/2$: if $k>n/2$, replace it by $n-k$ (recall that $|\Dnka nka|=|\Dnka n{n-k}a|$ by Proposition \ref{proposition:symmetry}). One can show that $B_{R,n,k}$ is minimal for the following choice of $R=R_{n,k}$:
    \[
        R_{n,k}=\begin{cases}
            \frac{n-2k-\sqrt{2(n-2k)^2-n^2}}{2(n-k)}&\text{if $k\leq\frac{2-\sqrt 2}{4}\cdot n$}\\
            \sqrt{\frac{k}{n-k}}&\text{if $k\geq\frac{2-\sqrt 2}{4}\cdot n$}
        \end{cases}.
    \]
    Figure \ref{fig:cauchy_estimate} compares for $n=1\,000$ the two upper bounds we have for $|\Dnka nka|$ on a logarithmic scale: the binomial coefficient $\binom nk$ in red, and the bound $B_{R,n,k}$ with the optimally chosen $R=R_{n,k}$ in blue. The $x$-axis represents the values for $k$, while the $y$-axis represents the binary logarithm of the two bounds. In case $|k-n/2|\leq\sqrt{2}/4\cdot n$, it can be observed that the Cauchy estimate yields particularly good bounds compared to the binomial coefficients. In fact, in this case, we get:
    \[
        |\Dnka nka|\leq 2^{n/4}\sqrt{\frac{n^n}{(n-k)^{n-k}\cdot k^k}}.
    \]
    \begin{figure}
        \centering
        \includegraphics[width=7cm]{cauchy_estimate.pdf}
        \caption{Cauchy estimate in blue vs.\ binomial coefficients in red}
        \label{fig:cauchy_estimate}
    \end{figure}
\end{remark}}

\subsection{Bounding the Walsh transform of $f$ experimentally}

We have just deduced that $|\wt f(a)|\leq 1+n\cdot 2^{3n/4}$ for every Boolean vector $a\in\F_2^n$. This is a pessimistic bound. In this section, we will discuss how a tighter bound on $|\wt f(a)|$ can be obtained in polynomial time in $n$.

@@

Consider now the complexity involved in computing the values of $\mathsf{D}_{n,k}^{p,q,r}$ for all $0 \leq k \leq n$ and for all triplets $(p,q,r)$ where $p+q+r=n/2$. We claim that this computation has a complexity of $\mathcal{O}(n^3 \log n)$. Towards this, according to Proposition \ref{proposition:generating_fct}, it is sufficient to expand the polynomials $(-z^2 + 2z + 1)^p \cdot (-z^2 - 2z + 1)^q \cdot (z^2 + 1)^r$ for all triplets $(p,q,r)$ such that $p+q+r=n/2$.

@@

Figure \ref{fig:walsh_bound_comparison} compares our experimental bound $B_n$ to the theoretical bound $1+n\cdot 2^{3n/4}$ and the binomial coefficient $\binom{n}{n/2}$, on a logarithmic scale. The first two and the last one have an asymptotic slope of $3/4$ and $1$, respectively.

\begin{figure}
    \centering
    \includegraphics[width=9cm]{comparison_walsh_bound.pdf}
    \caption{On a logarithmic scale, our experimental bound $B_n$ in terms of $n$ in red, our theoretical bound $1+n\cdot 2^{3n/4}$ from Proposition \ref{cor:bound_walsh_cauchy} in green, and the bound $\binom{n}{n/2}$ in blue.}
    \label{fig:walsh_bound_comparison}
\end{figure}

@@

\begin{proposition}\label{proposition:dtn}
	Let $t, n\in \N$ with $n$ even and $t\le n/2$, the following holds on $d_{t,n}$ for all $k\in [0,n]$:
	\[\wtk{d_{t,n}}{k}(a)=\sum_{\ell=0}^{2t} \Dnka{2t}{\ell}{b} \kraw{k-\ell}{\w(c)}{n-2t},\]
	where $b$ denotes the first $2t$ elements of $a$ and $c$ the last $n-2t$ elements.
